---
title: "Vitamin A Supply" Fortification set at status quo
author: "Gareth Osman"
date: "2025-04-07"
output: html_document
Objective: 'Using HH member specific nutrient vulnerability thresholds to calculate prevalence of apparent inadequacy'
---

> Packages

```{r}
library(tidyverse)
```

>Data

```{r}
ihs5 <- read.csv(here::here('paper_VHE/vuln.group/hh_mod_g_vulnfinal.csv'))
ihs5 <- select(ihs5, -X)
ihs5 <- ihs5 %>% filter(!is.na(VITA_RAE))

hh.hme <- read.csv(here::here('paper_VHE/vuln.group/hh.eq.factors.csv'))
hh.hme <- select(hh.hme, -X)

```

1. MN supply/food item = multiply quantity of foods consumed with MN composition of food
#Vitamin A supply 

```{r}
ihs5$va.supply<- ihs5$g100_d_nep*ihs5$VITA_RAE

```

2. Food group analysis dummy data frames
```{r}
#Disaggregation by food groups
cereal <- ihs5 %>% filter(food.group=="Cereals")
veg <- ihs5 %>% filter(food.group=="Vegetables")
meat <- ihs5 %>% filter(food.group=="Meat")
fat <- ihs5 %>% filter(food.group=="Fats & oils")
leg <- ihs5 %>% filter(food.group=="Legumes & nuts")
fruit <- ihs5 %>% filter(food.group=="Fruits")
fish <- ihs5 %>% filter(food.group=="Fish")
sugary <- ihs5 %>% filter(food.group=="Sugary foods/beverages")
dairy <- ihs5 %>% filter(food.group=="Milk & milk products")
egg <- ihs5 %>% filter(food.group=="Eggs")
tuber <- ihs5 %>% filter(food.group=="Roots & tubers")
misc <- ihs5 %>% filter(food.group=="Miscellaneous")

```

3. HH MN supply = Aggregate MN supply from all food items 
```{r}
d.va <- ihs5 %>% group_by(HHID) %>% summarise(va.supply=sum(va.supply))

```

4. Aggregate va supply per food group
```{r}

cereal.va <- cereal %>% group_by(HHID) %>% summarise(va.cereal=sum(va.supply))
tuber.va <- tuber %>% group_by(HHID) %>% summarise(va.tuber=sum(va.supply))
leg.va <- leg %>% group_by(HHID) %>% summarise(va.leg=sum(va.supply))
veg.va <- veg %>% group_by(HHID) %>% summarise(va.veg=sum(va.supply))
meat.va <- meat %>% group_by(HHID) %>% summarise(va.meat=sum(va.supply))
fruit.va <- fruit %>% group_by(HHID) %>% summarise(va.fruit=sum(va.supply))
dairy.va <- dairy %>% group_by(HHID) %>% summarise(va.dairy=sum(va.supply))
fat.va <- fat %>% group_by(HHID) %>% summarise(va.fat=sum(va.supply))
misc.va <- misc %>% group_by(HHID) %>% summarise(va.misc=sum(va.supply))
egg.va <- egg %>% group_by(HHID) %>% summarise(va.egg=sum(va.supply))
sugary.va <- sugary %>% group_by(HHID) %>% summarise(va.sugary=sum(va.supply))
fish.va <- fish %>% group_by(HHID) %>% summarise(va.fish=sum(va.supply))

```

5. Merge
```{r}
#Merged DF with all dietary variables
ihs5.va<- merge(x=hh.hme, y=d.va , by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.va<- merge(x=ihs5.va, y=cereal.va, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.va<- merge(x=ihs5.va, y=tuber.va, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.va<- merge(x=ihs5.va, y=leg.va, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.va<- merge(x=ihs5.va, y=veg.va, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.va<- merge(x=ihs5.va, y=meat.va, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.va<- merge(x=ihs5.va, y=fruit.va, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.va<- merge(x=ihs5.va, y=dairy.va, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.va<- merge(x=ihs5.va, y=fat.va, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.va<- merge(x=ihs5.va, y=misc.va, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.va<- merge(x=ihs5.va, y=egg.va, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.va<- merge(x=ihs5.va, y=sugary.va, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)
ihs5.va<- merge(x=ihs5.va, y=fish.va, by.x='HHID', by.y='HHID', fill=-9999, all.x = TRUE)

```
> APPARENT INTAKE METRIC BROKEN UP INTO TWO CATEGORIES:

**CATEGORY 1: APPARENT INTAKE/AFE**

#1. Supply per AFE

```{r}
#MN supply/AFE
ihs5.va$va.afe <- ihs5.va$va.supply/ihs5.va$hh.afe
#AFE inadequacy
ihs5.va <- ihs5.va %>% mutate(va.afe.inad = ifelse(va.afe <490, "Inadequate", "Adequate"))
```

**CATEGORY 2: APPARENT INTAKE/VULN.GROUP**

#2. Supply per vulnerable group threshhold 

```{r}
#MN supply/AFE
ihs5.va$va.vuln <- ihs5.va$va.supply/ihs5.va$vulnerable_equiv_va

#Vulnerable inadequacy
ihs5.va <- ihs5.va %>% filter(!is.na(va.vuln))
ihs5.va$va.vuln.inad <- NA

for (i in 1:nrow(ihs5.va)) {
  if(ihs5.va$va.vuln[i] < ihs5.va$vulnerable_equiv_H_AR_va[i]){
    ihs5.va$va.vuln.inad[i] <- 'Inadequate'
  } else {
    ihs5.va$va.vuln.inad[i] <- 'Adequate'
  }
}

```

> DONE: archive

```{r}
write.csv(ihs5.va, here::here ("paper_VHE/vuln.group/supply.Mn/results/va.vuln.csv"), row.names = FALSE)
          
```
